\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2023-07-02 cubble.Rmd
introduce unfold and check_key

```{r cubble-fun2, echo = TRUE}
identical(face_spatial(face_temporal(cb_spatial)), cb_spatial)
identical(face_temporal(face_spatial(cb_temporal)), cb_temporal)
```

@@

@@

The difference in LGA naming between both data sets triggers cubble to issue a warning, alerting the user to this discrepancy.  The warning message suggests there are some differences between the LGA encoding used by Victoria government and ABS and prompts analysts to check the mismatch using \code{check_key()}:

@@

```{r echo = TRUE}
res_sp <- match_spatial(climate_vic, river, spatial_n_group = 10)
print(res_sp, n = 20)
```

@@

```{r echo = TRUE}
res_sp <- match_spatial(
  climate_vic, river, 
  spatial_n_group = 10, return_cubble = TRUE)
(res_sp <- res_sp[-c(5, 8)] |> bind_rows())
```

@@

```{r echo = TRUE}
(res_tm <- res_sp |> 
  match_temporal(
    data_id = type, match_id = group,
    temporal_by = c("prcp" = "Water_course_level")))
```

@@

```{r echo = TRUE}
res_tm <- res_sp |> 
  match_temporal(
    data_id = type, match_id = group,
    temporal_by = c("prcp" = "Water_course_level"),
    return_cubble = TRUE)
(res_tm <- res_tm |> bind_rows() |> filter(group %in% c(1, 7, 6, 9)))
```

>>>>

```{r cubble-fun2, echo = TRUE}
face_spatial(face_temporal(cb_spatial))
```

@@

To enable operations involve both spatial and temporal variables, the function \code{unfold} incorporates spatial variables into the temporal cubble. Below is an example to include the coordinate columns (\code{long} and \code{lat}) into \code{cb_temporal} to prepare the data for a glyph map transformation, which will be discussed in Section \ref{st_transformation}.

```{r echo = TRUE}
cb_temporal |> unfold(long, lat)
```

@@

The difference in LGA naming between both data sets triggers cubble to issue a warning, alerting the user to this discrepancy.  The warning message suggests there are some differences between the LGA encoding used by Victoria government and ABS and prompts analysts to check the mismatch using \code{check_key()}, which takes the same inputs as \code{make_cubble}, but returns a summary of key matches between the spatial and temporal input data.

@@

```{r echo = TRUE}
res_sp <- match_spatial(df1 = climate_vic, df2 = river, 
                        spatial_n_group = 10)
print(res_sp, n = 20)
```

@@

```{r echo = TRUE}
res_sp <- match_spatial(
  df1 = climate_vic, df2 = river, 
  spatial_n_group = 10, return_cubble = TRUE)
(res_sp <- res_sp[-c(5, 8)] |> bind_rows())
```

@@

```{r echo = TRUE}
(res_tm <- match_temporal(data = res_sp,
                          data_id = type, match_id = group,
                          temporal_by = c("prcp" = "Water_course_level")))
```

@@

```{r echo = TRUE}
res_tm <- match_temporal(data =  res_sp,
                         data_id = type, match_id = group,
                         temporal_by = c("prcp" = "Water_course_level"),
                         return_cubble = TRUE)
(res_tm <- res_tm |> bind_rows() |> filter(group %in% c(1, 7, 6, 9)))
```
